---
title: 'Bios 6301: Assignment 5'
author: "Yeji Ko"
output: pdf_document
---

*Due Thursday, 14 October, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

40 points total.

Submit a single knitr file (named `homework5.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework5.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**15 points**

A problem with the Newton-Raphson algorithm is that it needs the derivative $f'$. If the derivative is hard to compute or does not exist, then we can use the *secant method*, which only requires that the function $f$ is continuous.

Like the Newton-Raphson method, the **secant method** is based on a linear approximation to the function $f$. Suppose that $f$ has a root at $a$. For this method we assume that we have *two* current guesses, $x_0$ and $x_1$, for the value of $a$. We will think of $x_0$ as an older guess and we want to replace the pair $x_0$, $x_1$ by the pair $x_1$, $x_2$, where $x_2$ is a new guess.

To find a good new guess x2 we first draw the straight line from $(x_0,f(x_0))$ to $(x_1,f(x_1))$, which is called a secant of the curve $y = f(x)$. Like the tangent, the secant is a linear approximation of the behavior of $y = f(x)$, in the region of the points $x_0$ and $x_1$. As the new guess we will use the x-coordinate $x_2$ of the point at which the secant crosses the x-axis.

The general form of the recurrence equation for the secant method is:

$$x_{i+1} = x_i - f(x_i)\frac{x_i - x_{i-1}}{f(x_i) - f(x_{i-1})}$$

Notice that we no longer need to know $f'$ but in return we have to provide *two* initial points, $x_0$ and $x_1$.

**Write a function that implements the secant algorithm.** Validate your program by finding the root of the function $f(x) = \cos(x) - x$. Compare its performance with the Newton-Raphson method -- which is faster, and by how much?  For this example $f'(x) = -\sin(x) - 1$.

```{r}
secant <- function(x0, x1, f){
  x2 <- x1 - f(x1)*{(x1-x0)/(f(x1)-f(x0))}
  while(abs(x2 - x1) < 0.0001){
    x0 <- x1
    x1 <- x2 
  }
  return(x2)
}

secant(1,2, function(x) cos(x) - x)
system.time(secant(1,2, function(x) cos(x) - x))

newton<- function(x1,f,fdev){
  x2 <- x1 - (f(x1)/fdev(x1))
  while (abs(x2 - x1) < 0.0001) {
     x1 <- x2
    
  }
 return(x2)
}

newton(1, function(x){cos(x) - x}, function(x) {-sin(x)-1})
system.time(newton(1, function(x){cos(x) - x}, function(x) {-sin(x)-1}))

```

The results from the secant function and the Newton-Raphson function are very similar, 0.7650 and 0.7503 respectively. Their running time was also very close, but the Newton-Raphson function was faster by 0.002.


### Question 2 ###

**20 points**

The game of craps is played as follows (this is simplified). First, you roll two six-sided dice; let x be the sum of the dice on the first roll. If x = 7 or 11 you win, otherwise you keep rolling until either you get x again, in which case you also win, or until you get a 7 or 11, in which case you lose.

Write a program to simulate a game of craps. You can use the following snippet of code to simulate the roll of two (fair) dice:

```{r}
craps <- function(n, m, k){
  result <- c(rep(NA,100))
  for(i in 1:k){  
    x <- sum(ceiling(6*runif(2)))
    result <- x
    if(x==n || x==m){
      print(cat("Game", i, ":", "result is win"))
      i <- i + 1
    }
    else{
      newroll <- sum(ceiling(6*runif(2)))
      result <- c(result, newroll)
      while(newroll!=x && newroll!=n && newroll!=m){
          newroll <- sum(ceiling(6*runif(2)))
          result <- c(result, newroll)
      }
      if(newroll==x){
        print(cat("Game", i, ":", "result is win"))
        i <- i + 1
      }
      else if(newroll==n || newroll==m){
        print(cat("Game", i, ":", "result is lose"))
        i <- i + 1
      }
    }
    print(result)
  }
}
```

1. The instructor should be able to easily import and run your program (function), and obtain output that clearly shows how the game progressed.  Set the RNG seed with `set.seed(100)` and show the output of three games. (lucky 13 points)
```{r}
set.seed(100)
craps(7,11,3)
```

2. Find a seed that will win ten straight games.  Consider adding an argument to your function that disables output. Show the output of the ten games. (7 points)
```{r}
craps <- function(n, m, k){
  result <- c(rep(NA,100))
  game <- data.frame(seq(1:k),NA)
  for(i in 1:k){  
    x <- sum(ceiling(6*runif(2)))
    result <- x
    if(x==n || x==m){
      game[i,2] <- 1 # 1 for "win"
      i <- i + 1
    }
    else{
      newroll <- sum(ceiling(6*runif(2)))
      result <- c(result, newroll)
      while(newroll!=x && newroll!=n && newroll!=m){
          newroll <- sum(ceiling(6*runif(2)))
          result <- c(result, newroll)
      }
      if(newroll==x){
        game[i,2] <- 1
        i <- i + 1
      }
      else if(newroll==n || newroll==m){
        game[i,2] <- 0 # 0 for "lose"
        i <- i + 1
      }
    }
  }
  #print(game)
  sum(game[,2])
}
```

```{r}
for(n in 1:10000){
set.seed(n)
x <- craps(7,11,10)
  if(x ==10){
    print(cat("seed:", n, "10 straight games"))
  }
  else{
    n <- n+1
  }
}
  
set.seed(880)
craps(7,11,10)
```

### Question 3 ###

**5 points**

This code makes a list of all functions in the base package:

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
length(funs)
```

Using this list, write code to answer these questions.

1. Which function has the most arguments? (3 points)

```{r}
x <- data.frame(seq(1:length(funs)), NA)
names(x) <- c("num", "nargs")
formals(unlist(funs[100])[[1]]) # tryout
length(formals(unlist(funs[100])[[1]]))

for(i in 1:length(funs)){
  x[i,2] <- length(formals(unlist(funs[i])[[1]]))
   i <- i+1
}

head(x,10)

x$num[which(x$nargs== max(x$nargs))]
funs[960] # a function named 'scan' has 22 arguments, which is the largest number of arguments.
```

2. How many functions have no arguments? (2 points)
```{r}
sum(x$nargs ==0) # 225 functions have no arguments. 
```
Hint: find a function that returns the arguments for a given function.
