---
title: 'Bios 6301: Assignment 2'
author: "Yeji Ko"
output:
  pdf_document: default
  html_document: default
---

*Due Tuesday, 21 September, 1:00 PM*

50 points total.

Add your name as `author` to the file's metadata section.

Submit a single knitr file (named `homework2.rmd`) by email to michael.l.williams@vanderbilt.edu.
Place your R code in between the appropriate chunks for each question.
Check your output by using the `Knit HTML` button in RStudio.

1. **Working with data** In the `datasets` folder on the course GitHub repo, you will find a file called `cancer.csv`, which is a dataset in comma-separated values (csv) format. This is a large cancer incidence dataset that summarizes the incidence of different cancers for various subgroups. (18 points)

    1. Load the data set into R and make it a data frame called `cancer.df`. (2 points)
    
    2. Determine the number of rows and columns in the data frame. (2)
    
    3. Extract the names of the columns in `cancer.df`. (2)
    
    4. Report the value of the 3000th row in column 6. (2)
    
    5. Report the contents of the 172nd row. (2)
    
    6. Create a new column that is the incidence *rate* (per 100,000) for each row.
    
    The incidence rate is the `(number of cases)/(population at risk)`, which in this case means 
    
    `(number of cases)/(population at risk) * 100,000`. (3)
    
    7. How many subgroups (rows) have a zero incidence rate? (2)
    
    8. Find the subgroup with the highest incidence rate.(3)
    
```{r}
# 1
cancer.df <- read.csv('/Users/yejiko/Downloads/cancer.csv')

# 2
nrow(cancer.df)
ncol(cancer.df)

# 3
colnames(cancer.df)

# 4
cancer.df[3000,6]

# 5
cancer.df[172,]

# 6

cancer.df[,'rate'] <- (cancer.df[,'incidence'] * 100000)/cancer.df[,'population'] 

# 7: 23191 rows have a zero incidence rate
nrow(cancer.df[which(cancer.df$rate==0),])

# 8
max(cancer.df$rate)
cancer.df[which(cancer.df$rate == max(cancer.df$rate)),]

```


2. **Data types** (10 points)

    1. Create the following vector: `x <- c("5","12","7")`. Which of the following commands will produce an error message? For each command, Either explain why they should be errors, or explain the non-erroneous result. (4 points)

            max(x)
            sort(x)
            sum(x)

    2. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            y <- c("5",7,12)
            y[2] + y[3]

    3. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            z <- data.frame(z1="5",z2=7,z3=12)
            z[1,2] + z[1,3]

```{r}
# 1
x <- c("5","12","7")
max(x) 
# there is no error but the maximum value is not "7". 
# Even if we change the order of values in the vector, it still outputs "7" 
# because each string represents some numeric values. 
# However, we cannot understand what is behind the scene intuitively. 

sort(x) # there is no error but this output is also incorrect. 
# We cannot understand what is behind the scene intuitively.

#sum(x) 
# it produces an error message 
# because it cannot make a summation of character values

# 2
y <- c("5",7,12)
#y[2] +y[3] 
str(y) 
# vector always contains the same type of values. 
# Because the last two elements have changed to character, this produce errors.

# 3
z <- data.frame(z1="5",z2=7,z3=12)
z[1,2] + z[1,3]
str(z) # dataframe can contain different data types, so this does not produce errors. 

```

3. **Data structures** Give R expressions that return the following matrices and vectors (*i.e.* do not construct them manually). (3 points each, 12 total)

    1. $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$
    
    2. $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$

    3. $\begin{pmatrix}
      0 & 1 & 1  \\
      1 & 0 & 1  \\
      1 & 1 & 0  \\
    \end{pmatrix}$

    4. $\begin{pmatrix}
      1 & 2 & 3 & 4 \\
      1 & 4 & 9 & 16 \\
      1 & 8 & 27 & 64  \\
      1 & 16 & 81 & 256 \\
      1 & 32 & 243 & 1024  \\
    \end{pmatrix}$

```{r}
# 1
c(seq(1:8), seq(7,1))

# 2
rep(1:5, times = 1:5)

# 3
mat <- matrix(1,3,3)
diag(mat) <- 0
mat

# 4
mat <- matrix(NA,5,4)
for(i in 1:5){
  mat[i,] <- (seq(1:4)^i)
}
mat
```

4. **Basic programming** (10 points)

    1. Let $h(x,n)=1+x+x^2+\ldots+x^n = \sum_{i=0}^n x^i$. Write an R program to calculate $h(x,n)$ using a `for` loop. As an example, use `x = 5` and `n = 2`. (5 points)
```{r}
x <- 5
n <- 2
sum <- 0
for(i in 0:n){
  sum <- sum + x^i
}
sum
```

    1. If we list all the natural numbers below 10 that are multiples of 3 or 5, we get 3, 5, 6 and 9. The sum of these multiples is 23. Write an R program to perform the following calculations. (5 points)

        1. Find the sum of all the multiples of 3 or 5 below 1,000. (3, [euler1])

        2. Find the sum of all the multiples of 4 or 7 below 1,000,000. (2)
        
```{r}
# 1
sum(c(seq(3,1000, by = 3), seq(5,1000, by = 5)))
table(seq(3,1000, by = 3)%%3==0) # all multiples of 3
table(seq(5,1000, by = 5)%%5==0) # all multiples of 3

# 2
sum(c(seq(4,1000000, by = 4), seq(7,1000000, by = 7)))
```

    1. Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with 1 and 2, the first 10 terms will be $(1, 2, 3, 5, 8, 13, 21, 34, 55, 89)$. Write an R program to calculate the sum of the first 15 even-valued terms. (5 bonus points, [euler2])

```{r}
fib <- c(1, 2)
i <- 2
sum <- 0
while (length(fib[fib %% 2 == 0]) < 15) {
  fib <- c(fib, fib[i-1] + fib[i])
  i <- i + 1
  if(fib[i]%%2 ==0){
    sum <- sum + fib[i]
  }
}

length(fib[fib %% 2 == 0])
sum
```
Some problems taken or inspired by projecteuler.

[euler1]: https://projecteuler.net/problem=1
[euler2]: https://projecteuler.net/problem=2
