---
title: 'Bios 6301: Assignment 3'
author: "Yeji Ko"
output:
pdf_document: default
html_document: default
---

*Due Tuesday, 28 September, 1:00 PM*

50 points total.

Add your name as `author` to the file's metadata section.

Submit a single knitr file (named `homework3.rmd`) by email to michael.l.williams@vanderbilt.edu.
Place your R code in between the appropriate chunks for each question.
Check your output by using the `Knit HTML` button in RStudio.

$5^{n=day}$ points taken off for each day late.

### Question 1 ###

**15 points**

Write a simulation to calculate the power for the following study
design.  The study has two variables, treatment group and outcome.
There are two treatment groups (0, 1) and they should be assigned
randomly with equal probability.  The outcome should be a random normal
variable with a mean of 60 and standard deviation of 20.  If a patient
is in the treatment group, add 5 to the outcome.  5 is the true
treatment effect.  Create a linear model for the outcome by the
treatment group, and extract the p-value (hint: see assigment1).
Test if the p-value is less than or equal to the alpha level, which
should be set to 0.05.

Repeat this procedure 1000 times. The power is calculated by finding
the percentage of times the p-value is less than or equal to the alpha
level.  Use the `set.seed` command so that the professor can reproduce
your results.

1. Find the power when the sample size is 100 patients. (10 points)
```{r}
treatment <- rbinom(100,1, 0.5)
outcome <- rnorm(100,60,20)
x <- data.frame(treatment, outcome)
x$outcome <- ifelse(x$treatment == 1, x$outcome + 5, x$outcome) # add 5 if treatment == 1
mod <- lm(outcome ~ treatment, data = x) # create a linear model 
coef(summary(mod))[2,4] # extract p-value

# repeat 1000 times
set.seed(1234)
mean(replicate(1e3, {
treatment <- rbinom(100,1, 0.5)
outcome <- rnorm(100,60,20)
x <- data.frame(treatment, outcome)
x$outcome <- ifelse(x$treatment == 1, x$outcome + 5, x$outcome) # add 5 if treatment == 1
mod <- lm(outcome ~ treatment, data = x) # create a linear model 
coef(summary(mod))[2,4]
}) < 0.05)

```

1. Find the power when the sample size is 1000 patients. (5 points)
```{r}
set.seed(1234)
mean(replicate(1e3, {
treatment <- rbinom(1000,1, 0.5)
outcome <- rnorm(1000,60,20)
x <- data.frame(treatment, outcome)
x$outcome <- ifelse(x$treatment == 1, x$outcome + 5, x$outcome) # add 5 if treatment == 1
mod <- lm(outcome ~ treatment, data = x) # create a linear model 
coef(summary(mod))[2,4]
}) < 0.05)

```

### Question 2 ###

**14 points**

Obtain a copy of the [football-values lecture](https://github.com/couthcommander/football-values).
Save the `2021/proj_wr21.csv` file in your working directory.  Read
in the data set and remove the first two columns.

1. Show the correlation matrix of this data set. (4 points)
```{r}
proj_wr21 <- read.csv('proj_wr21.csv')
x <- proj_wr21[-c(1:2)]
cor(x)
```

1. Generate a data set with 30 rows that has a similar correlation
structure.  Repeat the procedure 1,000 times and return the mean
correlation matrix. (10 points)
```{r}
(rho=cor(x))
(vcov=var(x))		
(means=colMeans(x))

require(MASS)
sim = mvrnorm(30, mu = means, Sigma = vcov)
sim = as.data.frame(sim)

(rho.sim=cor(sim))   # create a simulated correlation matrix
rho                  # compare with the original correlation matrix

keep.1 = 0
loops=1000

for (i in 1:loops) {
      sim = mvrnorm(30, mu = means, Sigma = vcov)
      keep.1=keep.1+ cor(sim)/loops   
}

rho
keep.1  
```

### Question 3 ###

**21 points**

Here's some code:

```{r}
nDist <- function(n = 100) {
    df <- 10
    prob <- 1/3
    shape <- 1
    size <- 16
    list(
        beta = rbeta(n, shape1 = 5, shape2 = 45),
        binomial = rbinom(n, size, prob),
        chisquared = rchisq(n, df),
        exponential = rexp(n),
        f = rf(n, df1 = 11, df2 = 17),
        gamma = rgamma(n, shape),
        geometric = rgeom(n, prob),
        hypergeometric = rhyper(n, m = 50, n = 100, k = 8),
        lognormal = rlnorm(n),
        negbinomial = rnbinom(n, size, prob),
        normal = rnorm(n),
        poisson = rpois(n, lambda = 25),
        t = rt(n, df),
        uniform = runif(n),
        weibull = rweibull(n, shape)
    )
}

```

1. What does this do? (3 points)

    ```{r}
    round(sapply(nDist(500), mean), 2)
    
    ```
    
    ```
    This prints out the means of 500-times-replicated test statistics 
    from 15 different distributions, whch is rounded with two decimal places. 
    ```

1. What about this? (3 points)

    ```{r}
    sort(apply(replicate(20, round(sapply(nDist(10000), mean), 2)), 1, sd))
    ```
    
    ```
    This replicates the means of 10000-times-replicated test statistics, 
    each from 15 different distributions, rounded with two decimal places, 20 times. 
    Then it calculates the standard deviation for each distribution and orders from lowest to highest. 
    ```

    In the output above, a small value would indicate that `N=10,000` 
    would provide a sufficent sample size as to estimate the mean of the distribution. 
    Let's say that a value *less than 0.02* is "close enough".

1. For each distribution, estimate the sample size required to simulate the distribution's mean. (15 points)

Don't worry about being exact. It should already be clear that N < 10,000 for many of the distributions. You don't have to show your work. Put your answer to the right of the vertical bars (`|`) below.

distribution|N
---|---
beta|16
binomial|21
chisquared|32
exponential|34
f|35
gamma|41
geometric|53
hypergeometric|68
lognormal|76
negbinomial|83
normal|85
poisson|102
t|107
uniform|110
weibull|119

    ```{r}
n <- 10
i <- 1
size <- c()
set.seed(1234)
for(i in 1:15){repeat{
    truth <- data.frame(matrix(rep(round(sapply(nDist(10000), mean), 2),n), ncol = 15, byrow = TRUE))
    x <- nDist(n)
    p <- t.test(x[[i]],truth[[i]], alternative='two.sided')$p.value
  if(p < 0.05) break
  n <- n + 1
  size[i] <- n
}
  i <- i + 1
}

size
    ```
