---
title: 'Bios 6301: Assignment 8'
author: 'Yeji Ko'
output: pdf_document
---

*Due Tuesday, 16 November, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

30 points total.

Submit a single knitr file (named `homework8.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework8.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**15 points**

Install the `readxl` package and run the following

```{r}
fn <- 'icd10.xlsx'
if(file.access(fn, mode = 4) == -1) {
    url <- "https://www.cdc.gov/nhsn/xls/icd10-pcs-pcm-nhsn-opc.xlsx"
    download.file(url, destfile = fn, mode = 'wb')
}
dat <- readxl::read_excel(fn, sheet = 2)
```

1. Show the class of `dat`. (1 point)

```{r}
class(dat)
```

2. Show the methods available for objects of the given class (if there are multiple classes, show methods for all classes). (3 points)

```{r}
for(i in 1:length(class(dat))){
  print(methods(class=class(dat)[i]))
}

```

3. If you call `print(dat)`, what print method is being dispatched? (1 point)

```{r}
findMethod <- function(generic, ...) {
  ch <- deparse(substitute(generic))
  f <- X <- function(x, ...) UseMethod("X")
  for(m in methods(ch)) assign(sub(ch, "X", m, fixed = TRUE), "body<-"(f, value = m))
  X(...)
}

findMethod(print, dat)
```

4. Set the class of `dat` to be a data.frame. (1 point)

```{r}
class(dat) <- "data.frame"
```

5. If you call `print(dat)` again, what print method is being dispatched? (1 point)

```{r}
findMethod(print, dat)
```

Define a new generic function `nUnique` with the code below.

```{r}
nUnique <- function(x) {
    UseMethod('nUnique')
}
```

6. Write a default method for `nUnique` to count the number of unique values in an element. (2 points)

```{r}
nUnique.default <- function(x) {
  length(unique(x))
}
```

7. Check your function (2 points)

```{r}
nUnique(letters) # should return 26
nUnique(sample(10, 100, replace = TRUE)) # should return 10 (probably)
```

8. Write a data.frame method for `nUnique` to operate on data.frame objects.
This version should return counts for each column in a data.frame. (2 points)

```{r}
nUnique.data.frame <- function(x) {
data <- c()

for(i in 1: ncol(x)){
  data <-c(data, length(sapply(x, unique)[[i]]))
}
names(data) <- colnames(x)
print(data)
}

```

9. Check your function (2 points)

```{r}
nUnique(dat)
```

### Question 2 ###

**15 points**

Programming with classes.  The following function will generate random patient information.

```{r}
makePatient <- function() {
  vowel <- grep("[aeiou]", letters)
  cons <- grep("[^aeiou]", letters)
  name <- paste(sample(LETTERS[cons], 1), sample(letters[vowel], 1), sample(letters[cons], 1), sep='')
  gender <- factor(sample(0:1, 1), levels=0:1, labels=c('female','male'))
  dob <- as.Date(sample(7500, 1), origin="1970-01-01")
  n <- sample(6, 1)
  doa <- as.Date(sample(1500, n), origin="2010-01-01")
  pulse <- round(rnorm(n, 80, 10))
  temp <- round(rnorm(n, 98.4, 0.3), 2)
  fluid <- round(runif(n), 2)
  list(name, gender, dob, doa, pulse, temp, fluid)
}
```

1. Create an S3 class `medicalRecord` for objects that are a list with the named elements `name`, `gender`, `date_of_birth`, `date_of_admission`, `pulse`, `temperature`, `fluid_intake`. Note that an individual patient may have multiple measurements for some measurements.  Set the RNG seed to `8` and create a medical record by taking the output of `makePatient`.  Print the medical record, and print the class of the medical record. (5 points)

```{r}
set.seed(8)
patient <- makePatient()

names(patient) <- c("name", "gender", "date_of_birth", "date_of_admission", 
                    "pulse", "temperature", "fluid_intake")
class(patient) <- 'medicalRecord'
print(patient)
attributes(patient)
```

2. Write a `medicalRecord` method for the generic function `mean`, which returns averages for pulse, temperature and fluids. Also write a `medicalRecord` method for `print`, which employs some nice formatting, perhaps arranging measurements by date, and `plot`, that generates a composite plot of measurements over time.  Call each function for the medical record created in part 1. (5 points)

```{r}
library(dplyr)

mean.medicalRecord <- function(x) {
  vector <- c(mean(x$pulse), mean(x$temperature), mean(x$fluid_intake))
  names(vector) <- c("pulse", "temperature", "fluid_intake")
  print(vector)
}

print.medicalRecord <- function(x) {
  data <- data.frame(matrix(NA, length(x$date_of_admission), length(x)))
  for(i in 1:length(x)){
  data[,i] <- x[[i]]
  }
  colnames(data) <- names(x)
  data <- arrange(data,date_of_admission)
  data
}

plot.medicalRecord <- function(x) {
  par(mfrow = c(1,3))
  plot(x$date_of_admission, x$pulse, col = 'red')
  plot(x$date_of_admission, x$temperature, col = 'red')
  plot(x$date_of_admission, x$fluid_intake, col = 'red')
}

mean(patient)
print(patient)
plot(patient)
```

3. Create a further class for a cohort (group) of patients, and write methods for `mean` and `print` which, when applied to a cohort, apply mean or print to each patient contained in the cohort. Hint: think of this as a "container" for patients.  Reset the RNG seed to 8 and create a cohort of ten patients, then show the output for `mean` and `print`. (5 points)

```{r}
set.seed(8)

cohort <- list()

for(i in 1:10) {
patient <- makePatient()
names(patient) <-  c("name", "gender", "date_of_birth", "date_of_admission", 
                       "pulse", "temperature", "fluid_intake")
# class(patient) <- "medicalRecord"
cohort[[i]] <- patient
}

class(cohort) <- "medicalCohort"

mean.medicalCohort <- function(x){
  mat <- matrix(NA, 10, 3)
  for(i in 1:10) {
    mat[i,] <- mean.medicalRecord(x[[i]])
  }
  colMeans(mat)
}

mean(cohort)   

print.medicalCohort <- function(x) {
  dat <- data.frame()
  for(i in 1:10) {
    dat <- rbind(dat, print.medicalRecord(cohort[[i]]))
  }
  dat
}

print(cohort)
```